### 2022-12-05 dissertation/chapters/04-hyper-heuristics.tex
✨ (review): Spelling and grammar, derived publications fix

A recent suggestion related to the field of \index{meta-learning}meta-learning is to dynamically select and/or adjust the \index{heuristic}heuristic characteristics and heaviours used throughout the search process and not just the \index{heuristic}heuristic hyper-parameters. This approach focuses on the hybridisation of \index{heuristic}heuristic paradigms. By dynamically combining the best of different \index{heuristic} paradigms throughout the learning process, a trade-off can be made between focussing on a particular solution (exploitation), in comparison to seeking out novel solutions (exploration) during the search process. The dynamic trade-off between exploration and exploitation could also be seen as a dynamic trade-off between the strengths and weaknesses of different \index{heuristic}heuristics during the search process.

>>>>

A recent suggestion related to the field of \index{meta-learning}meta-learning is to dynamically select and/or adjust the \index{heuristic}heuristic characteristics and behaviours used throughout the search process and not just the \index{heuristic}heuristic hyper-parameters. This approach focuses on the hybridisation of \index{heuristic}heuristic paradigms. By dynamically combining the best of different \index{heuristic} paradigms throughout the learning process, a trade-off can be made between focussing on a particular solution (exploitation), in comparison to seeking out novel solutions (exploration) during the search process. The dynamic trade-off between exploration and exploitation could also be seen as a dynamic trade-off between the strengths and weaknesses of different \index{heuristic}heuristics during the search process.

### 2022-12-05 dissertation/chapters/05-probability.tex
💚 Dedicated to Taylah, Beané and Aleah
✨ (review): Spelling and grammar, derived publications fix

Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of of Chance}; a book that is widely regarded as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of \index{Bayes' theorem}Bayes' theorem was introduced.

@@

In everyday conversation, the term \textit{probability} is a measure of belief in the occurrence of a future event~\cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that cannot be predicted with absolute certainty~\cite{ref:wackerly:2014}. Probability can be be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event, $A$, might occur is denoted by $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials, is often remarkably stable.

@@

\begin{figure}[htb]
      \centering
      \includegraphics[width=0.9\textwidth]{images/dirichlet.pdf}
      \caption{The \acfp{PDF} for the \index{Dirichlet probability distribution}Dirichlet probability distribution over the 2-simplex. The concentration parameters, $\boldsymbol{\alpha}$, are varied. The values of the \acs{PDF} are shown by the color maps with contour lines at equal values as indicated in the color bars~\cite{ref:dirichlet:2020}.}
      \label{fig:probability:probability_distributions:dirichlet}
\end{figure}

@@

\begin{figure}[htbp]
      \centering
      \begin{subfigure}{0.49\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/coin_flip_samples_100.pdf}
            \caption{100 Samples}
            \label{fig:probability:probability_distributions:bernoulli:coin_100}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/coin_flip_samples_10000.pdf}
            \caption{10000 Samples}
            \label{fig:probability:probability_distributions:bernoulli:coin_10000}
      \end{subfigure}
      \par\bigskip
      \caption{An illustration of the the coin-flip simulation for different sample sizes that show the convergence of the mean as per the \acf{CLT}.}
      \label{fig:probability:probability_distributions:bernoulli:coin}
\end{figure}

@@

\subsection{Multinomial Distribution}
\label{sec:probability:probability_distributions:multinomial}

>>>>

Probability theory and statistics can be traced back to as early as the 18th century. In 1718,~\citeauthor{ref:demoivre:1718}~\cite{ref:demoivre:1718} published the \textit{Doctrine of Chance}; a book that is widely regarded as the first published book on probability theory. In 1763, Thomas Bayes~\cite{ref:bayes:1763} published an article titled \textit{An Essay towards solving a Problem in the Doctrine of Chances} where the first version of \index{Bayes' theorem}Bayes' theorem was introduced.

@@

In everyday conversation, the term \textit{probability} is a measure of belief in the occurrence of a future event~\cite{ref:wackerly:2014}. Probability is a necessary tool used in many fields including physics, biology, chemistry and computer science. These fields contain many cases that generate observations that cannot be predicted with absolute certainty~\cite{ref:wackerly:2014}. Probability can be inferred and confirmed through past events. These events are referred to as \textit{random} or \textit{stochastic} events. The probability that a certain event, $A$, might occur is denoted by $P(A)$. Although these random events cannot be predicted with absolute certainty, the relative frequency with which they occur over many trials, is often remarkably stable.

@@

\begin{figure}[htb]
      \centering
      \includegraphics[width=0.9\textwidth]{images/dirichlet.pdf}
      \caption{The \acfp{PDF} for the \index{Dirichlet probability distribution}Dirichlet probability distribution over the 2-simplex. The concentration parameters, $\boldsymbol{\alpha}$, are varied. The values of the \acs{PDF} are shown by the colour maps with contour lines at equal values as indicated in the colour bars~\cite{ref:dirichlet:2020}.}
      \label{fig:probability:probability_distributions:dirichlet}
\end{figure}

@@

\begin{figure}[htbp]
      \centering
      \begin{subfigure}{0.49\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/coin_flip_samples_100.pdf}
            \caption{100 Samples}
            \label{fig:probability:probability_distributions:bernoulli:coin_100}
      \end{subfigure}
      \begin{subfigure}{0.49\textwidth}
            \centering
            \includegraphics[width=\textwidth]{images/coin_flip_samples_10000.pdf}
            \caption{10000 Samples}
            \label{fig:probability:probability_distributions:bernoulli:coin_10000}
      \end{subfigure}
      \par\bigskip
      \caption{An illustration of the coin-flip simulation for different sample sizes that show the convergence of the mean as per the \acf{CLT}.}
      \label{fig:probability:probability_distributions:bernoulli:coin}
\end{figure}

@@

\subsection{Multinomial Probability Distribution}
\label{sec:probability:probability_distributions:multinomial}

### 2022-12-05 dissertation/chapters/06-bayesian-hyper-heuristic.tex
✨ (review): Spelling and grammar, derived publications fix

      \item \textbf{position delta}: The last computed position delta between the current timestep and the previous timestep.

@@

      \item \textbf{loss}: A parameter that tracks the loss as achieved by the the entity throughout training.
\end{itemize}

@@

The \acs{BHH} incorporates a form of probabilistic modeling in its selection mechanism. Probability is calculated based on \index{heuristic}heuristic-entity performance over time. Evidence of \index{heuristic}heuristic-entity performance is thus required for the \acs{BHH} to learn. Since the performance log can become very big, only a sliding window of the performance history is maintained at each step in the learning process. The sliding window is also referred to as a \textit{replay} buffer, a term borrowed from the field of \acf{RL}. The replay window size is defined as a hyper-parameter of the \acs{BHH}. The performance log is then simply a table of events and metrics. The \acs{BHH} tracks the following metrics in the performance log:

@@

      \item \textbf{symmetric}: Credit is assigned to all entity-\index{heuristic}heuristic combinations, regardless of their performance. The symmetric credit assignment strategy does not randomly assign credit, but rather assigns credit to all events. In effect, no learning and performance-bias is achieved with the symmetric credit assignment strategy and thus the symmetric credit assignment strategy is implemented as a basis for comparison. Comparing a credit assignment staretegy to the symmetric credit assignment indicates the ability of the other credit assignment strategy to have an effect on the learning process and performance of the \acs{BHH}.
\end{itemize}

@@

\begin{itemize}
      \item $I$ denote the maximum number of instances in the replay window.
      \item $J$ denote the \index{entity pool}entity pool (population) size.
      \item $K$ denote the \index{heuristic pool}heuristic pool size.
      \item $L$ denote the number of credit assignment output classes. Since the output of credit assignment is boolean, $L = 2$.

@@

This section aims to dissect the probabilistic model that is presented in Equestion~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to}. The \acs{BHH} implements a form of Naïve Bayes classifier, and thus independence between events can be assumed. The following derived \acp{PMF} are provided as fundamental building blocks to show the mechanism by which the \acs{BHH} learns.

@@

Equations~\eqref{eq:bhh:selection_mechanism:naive_bayes:h_pmf}, \eqref{eq:bhh:selection_mechanism:naive_bayes:EgH_pmf} and \eqref{eq:bhh:selection_mechanism:naive_bayes:CgH_pmf} can be substituded into the proportional evaluation of the predictive model as given in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to}, resulting in

@@

When Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} is evaluated, the numerical stability is shown to underflow if the resulting probabilities from its parts are very small. Multiplication of multiple fractional parameters leads to an even smaller fractional number. Probabilities might be very low at some points during training. Consider an example where training has stagnated, effectively leading to a scenario where a credit assignment strategy never fulfills a credit assignment, yielding an extremely small probability for $\boldsymbol{\psi}$. A solution to the aforementioned problem is to apply the \textit{log-sum-exp} trick. The transformation of Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} using the log-sum-exp trick is given as

@@

This chapter provided extensive detail on the inner workings and design of the \acs{BHH}. The \acs{BHH} was formally classified and the details around various components of the \acs{BHH}'s architecture was presented. Formal mathematical descriptions of the Bayesian selection method have been provided. The selection mechanism was presented as a probabilistic, predictive classifier. The optimisation process, by means of Bayesian analysis, has also been presented. Hyper-parameters were discussed in detail and a pseudo code implementation of the \acs{BHH} was presented.

>>>>

      \item \textbf{position delta}: The last computed position delta between the current time step and the previous time step.

@@

      \item \textbf{loss}: A parameter that tracks the loss as achieved by the entity throughout training.
\end{itemize}

@@

The \acs{BHH} incorporates a form of probabilistic modelling in its selection mechanism. Probability is calculated based on \index{heuristic}heuristic-entity performance over time. Evidence of \index{heuristic}heuristic-entity performance is thus required for the \acs{BHH} to learn. Since the performance log can become very big, only a sliding window of the performance history is maintained at each step in the learning process. The sliding window is also referred to as a \textit{replay} buffer, a term borrowed from the field of \acf{RL}. The replay window size is defined as a hyper-parameter of the \acs{BHH}. The performance log is then simply a table of events and metrics. The \acs{BHH} tracks the following metrics in the performance log:

@@

      \item \textbf{symmetric}: Credit is assigned to all entity-\index{heuristic}heuristic combinations, regardless of their performance. The symmetric credit assignment strategy does not randomly assign credit, but rather assigns credit to all events. In effect, no learning and performance-bias is achieved with the symmetric credit assignment strategy and thus the symmetric credit assignment strategy is implemented as a basis for comparison. Comparing a credit assignment strategy to the symmetric credit assignment indicates the ability of the other credit assignment strategy to have an effect on the learning process and performance of the \acs{BHH}.
\end{itemize}

@@

\begin{itemize}
      \item $I$ denote the maximum number of instances in the replay window.
      \item $J$ denote the \index{entity pool}entity pool (population) size.
      \item $K$ denote the \index{heuristic pool}heuristic pool size.
      \item $L$ denote the number of credit assignment output classes. Since the output of credit assignment is Boolean, $L = 2$.

@@

This section aims to dissect the probabilistic model that is presented in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to}. The \acs{BHH} implements a form of Naïve Bayes classifier, and thus independence between events can be assumed. The following derived \acp{PMF} are provided as fundamental building blocks to show the mechanism by which the \acs{BHH} learns.

@@

Equations~\eqref{eq:bhh:selection_mechanism:naive_bayes:h_pmf}, \eqref{eq:bhh:selection_mechanism:naive_bayes:EgH_pmf} and \eqref{eq:bhh:selection_mechanism:naive_bayes:CgH_pmf} can be substituted into the proportional evaluation of the predictive model as given in Equation~\eqref{eq:bhh:selection_mechanism:predictive_model_prop_to}, resulting in

@@

When Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} is evaluated, the numerical stability is shown to underflow if the resulting probabilities from its parts are very small. Multiplication of multiple fractional parameters leads to an even smaller fractional number. Probabilities might be very low at some points during training. Consider an example where training has stagnated, effectively leading to a scenario where a credit assignment strategy never fulfils a credit assignment, yielding an extremely small probability for $\boldsymbol{\psi}$. A solution to the aforementioned problem is to apply the \textit{log-sum-exp} trick. The transformation of Equation~\eqref{eq:bhh:selection_mechanism:naive_bayes:HgEC_pmf} using the log-sum-exp trick is given as

@@

This chapter provided extensive detail on the inner workings and design of the \acs{BHH}. The \acs{BHH} was formally classified and the details around various components of the \acs{BHH}'s architecture was presented. Formal mathematical descriptions of the Bayesian selection method have been provided. The selection mechanism was presented as a probabilistic, predictive classifier. The optimisation process, by means of Bayesian analysis, has also been presented. Hyper-parameters were discussed in detail and a pseudo-code implementation of the \acs{BHH} was presented.

### 2022-12-05 dissertation/chapters/07-methodology.tex
✨ (review): Spelling and grammar, derived publications fix

\begin{table}[htb]
      \centering
      \caption{Regression datasets}
      \label{tab:methodology:datasets:regression}%
      \par\bigskip
      \resizebox{\textwidth}{!}{
            \begin{tabular}{cccccccc}
                  \textbf{dataset}    & \textbf{output}          & \textbf{types} & \textbf{attributes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}         \\
                  \midrule
                  fish toxicity       & multivariate             & real           & 7                   & 908                & 64             & 15             & ~\cite{ref:cassotti:2015} \\
                  housing             & univariate               & real           & 13                  & 506                & 32             & 16             & ~\cite{ref:harrison:1978} \\
                  forest fires        & multivariate             & real           & 13                  & 517                & 32             & 17             & ~\cite{ref:cortez:2007}   \\
                  student performance & multivariate             & integer        & 33                  & 649                & 32             & 21             & ~\cite{ref:cortez:2008}   \\
                  parkinsons          & multivariate             & integer, real  & 26                  & 5875               & 256            & 23             & ~\cite{ref:tsanas:2009}   \\
                  air quality         & multivariate, timeseries & real           & 15                  & 9358               & 256            & 37             & ~\cite{ref:de:2008}       \\
                  bike                & univariate               & integer, real  & 16                  & 17389              & 256            & 68             & ~\cite{ref:fanaee:2014}   \\
            \end{tabular}%
      }
\end{table}%

@@

The behavioural case study provides an implementation of the \acs{BHH} baseline which, by default, has a replay window size of 10. The baseline configuration is provided and compared to an implementation of a \acs{BHH} configuration where the replay window size is large (250), as well as an implementation of the \acs{BHH} using the \textit{symmetric} credit assignment strategy. The large replay window size configuration is to used to show the behaviour of the \acs{BHH} when it has access to a large number of performance log samples, which increases the statistical certainty of the learning outcome. The \acs{BHH} configuration that uses the \textit{symmetric} credit assignment strategy is used to illustrate the behaviour of the \acs{BHH} where no performance bias takes place and thus, no learning occurs.

>>>>

\begin{table}[htb]
      \centering
      \caption{Regression datasets}
      \label{tab:methodology:datasets:regression}%
      \par\bigskip
      \resizebox{\textwidth}{!}{
            \begin{tabular}{cccccccc}
                  \textbf{dataset}    & \textbf{output}           & \textbf{types} & \textbf{attributes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}         \\
                  \midrule
                  fish toxicity       & multivariate              & real           & 7                   & 908                & 64             & 15             & ~\cite{ref:cassotti:2015} \\
                  housing             & univariate                & real           & 13                  & 506                & 32             & 16             & ~\cite{ref:harrison:1978} \\
                  forest fires        & multivariate              & real           & 13                  & 517                & 32             & 17             & ~\cite{ref:cortez:2007}   \\
                  student performance & multivariate              & integer        & 33                  & 649                & 32             & 21             & ~\cite{ref:cortez:2008}   \\
                  parkinsons          & multivariate              & integer, real  & 26                  & 5875               & 256            & 23             & ~\cite{ref:tsanas:2009}   \\
                  air quality         & multivariate, time series & real           & 15                  & 9358               & 256            & 37             & ~\cite{ref:de:2008}       \\
                  bike                & univariate                & integer, real  & 16                  & 17389              & 256            & 68             & ~\cite{ref:fanaee:2014}   \\
            \end{tabular}%
      }
\end{table}%

@@

The behavioural case study provides an implementation of the \acs{BHH} baseline which, by default, has a replay window size of 10. The baseline configuration is provided and compared to an implementation of a \acs{BHH} configuration where the replay window size is large (250), as well as an implementation of the \acs{BHH} using the \textit{symmetric} credit assignment strategy. The large replay window size configuration is used to show the behaviour of the \acs{BHH} when it has access to a large number of performance log samples, which increases the statistical certainty of the learning outcome. The \acs{BHH} configuration that uses the \textit{symmetric} credit assignment strategy is used to illustrate the behaviour of the \acs{BHH} where no performance bias takes place and thus, no learning occurs.
